---
title: "Convenience functions for Fiji/ImageJ"
author: "John Minter"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Convenience functions for Fiji/ImageJ}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, cache = F, echo = F, message = F, warning = F, tidy = F, comment=NA}
library(knitr)
options(width = 72)
```
Fiji/ImageJ has a particle analysis function called `Analyze Particles...`.

To set the parameters in a macro we need to construct a string
that we will name `ap_str2`: This was split to two lines for easier
reading. It is one long line in the original... 

```
ap_str2 = "size=" + min_size + "-Infinity circularity=" + min_circ + "-1.00
show=Overlay display exclude add in_situ"
```

The `min_size` variable is implemented as a `String` as is the `min_circ`
variable. Our focus is on the `min-size` variable. It is the
minimum area in square microns to detect. Let's consider that we want to
compute the area if we wanted to detect particles greater than or equal
to 5 **nm** in diameter.

First, we would load the library

```{r compute_area, message=FALSE}
library(particlesizeR)
```


```{r comput_area, comment=NA}
print(ecd_nm_to_area_um(5.0))

```

To work backwards, if we wanted to know the diameter in nm from
an area cutoff of `7.854e-5` square microns, we would run

```{r compute_diameter, comment=NA}
print(area_um_to_ecd_nm(7.854e-5))
```


