---
title: "Test Gaussian Mixture"
author: "John Minter"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Test Gaussian Mixture}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

First, let's load the required packages.

```{r}
library(particlesizeR)
library(mixtools)
```

Now, let's generate a two Gaussian mixture. Because we consider this a particle size distribution, we need to require the values to be positive.

```{r}
samples <- gen.two.gaussian.mixture(10000, 0.5, 3.0, 1.0, 12.0, 2.0)
samples <- samples[samples > 0]
```

Let's now construct a histogram object. This will give us access to arrays of midpoints and density values.

```{r, fig.width=7}
o.mar <- par("mar")
par(mar=c(4.1,4.1,0,0.5))
his <- hist(samples, breaks=250, freq=FALSE, main="")
par(mar=o.mar)
```

Eventually, we want to plot the results. We will build a plot from base graphics because we have a lot of control. We will also generate a series of x (diameter) values to model.

```{r, fig.width=7}
o.mar <- par("mar")
par(mar=c(4.1,4.1,0,0.5))
x = seq(0,20,.1)
x1 <- min(samples)
x2 <- max(samples)
truth <- gen.two.gaussian.dist(0.5, 3.0, 1.0, 12.0, 2.0, 0, 20, 0.1)
ymax <- 1.1*max(max(his$density), max(truth[,2]))
plot(c(x1, x2), c(0., ymax), type='n', xlab='x', ylab='density')
points(his$mids, his$density, pch=19, cex=0.75)
lines(truth[, 1], truth[, 2],col="red",lwd=4)
par(mar=o.mar)
```

We want to compute the mixture model. We will make a guess at the parameters from the histogram.

```{r}
out <- normalmixEM(samples, lambda=0.5, mu=c( 2, 9.), sigma=c(.8, 2.))
```

We can now compute the fitted values:

```{r}
fit <- out$lambda[1]*dnorm(x, out$mu[1], out$sigma[1]) + 
       out$lambda[2]*dnorm(x, out$mu[2],out$sigma[2])
```

And make a final plot.


```{r, fig.width=7, fig.height=5}
o.mar <- par("mar")
par(mar=c(4.1,4.1,0.5,0.5))
plot(c(x1, x2), c(0., ymax), type='n', xlab='x', ylab='density')#, 
#    main="Density Estimate of the Gaussian Mixture Model")
points(his$mids, his$density, pch=19, cex=0.75)
lines(truth[, 1], truth[, 2],col="red",lwd=6)
lines(x, fit, col="blue", lwd=2)
legend("topright", c("True Density","Estimated Density", "fit"),
       col=c("black","red", "blue"), lwd=c(NA, 4, 2), pch=c(19,NA,NA))
par(mar=o.mar)
```
