---
title: "Test Gaussian Mixture"
author: "John Minter"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Test Gaussian Mixture}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

First, let's load the required packages.

```{r}
library(particlesizeR)
library(mixtools)
library(ggplot2)
```

Now, let's generate a two Gaussian mixture. Because we consider this a particle size distribution, we need to require the values to be positive.

```{r}
samples <- gen_two_gaussian_mixture(10000, 0.5, 3.0, 1.0, 12.0, 2.0)
samples <- samples[samples > 0]
```

Let's now construct a histogram object. This will give us access to arrays of midpoints and density values.

```{r first_plot, fig.width=7}
library(ggplot2)

df <- data.frame(samples=samples)
plt <- ggplot(df, aes(x=samples)) + geom_histogram(bins=250)
print(plt)
```

Let's look at what is in a ggplot2 histogram, starting at the beginning:

```{r examine_histo_head}
pg <- ggplot_build(plt)
head(pg$data[[1]])
```
Note that we have access to a vector of midpoints (`x`) as well as a vector
of lower boundaries (`xmin`) and a vector of upper boundaries (`xmax`) as
well as the `density` and the `ncount`. 


Next, look at the end of the data...

```{r examine_histo_tail}
tail(pg$data[[1]])
```

Eventually, we want to plot the results. We will build a plot using the newer
`ggplot2` package. We will also generate a series of x (diameter) values to
model.

First, we will generate the data

```{r generate_data}
#x = seq(0, 20, 0.1)
#x1 <- min(samples)
# x2 <- max(samples)
truth <- gen_two_gaussian_dist(0.5, 3.0, 1.0, 12.0, 2.0, 0, 20, 0.1)

print(c(min(truth), max(truth)))
print(head(truth))
```

```{r, fig.width=7}
df_work <- data.frame(diam=pg$data[[1]]$x, count=pg$data[[1]]$count)
plt <- ggplot(df_work, aes(diam,count)) + geom_point() +
       geom_line(data=truth, aes(x,800*dist), colour='darkred', size=2)+
       ylab("count") + # note the Unicode string for mu
       xlab("diameter [nm]") +
       ggtitle("Diameter distribution of two Gaussian distributions") +
       theme(axis.text=element_text(size=12), axis.title=element_text(size=14),
              plot.title = element_text(hjust = 0.5)) # center the title
print(plt)




```


We want to compute the mixture model. We will make a guess at the parameters from the histogram.

```{r mixModel}
out <- normalmixEM(samples, lambda=0.5, mu=c( 2, 9.), sigma=c(.8, 2.))

print("lambda")
print(out$lambda)
print("mu")
print(out$mu)
print("sigma")
print(out$sigma)
```

We can now compute the fitted values:

```{r}
x=100
fit <- out$lambda[1]*dnorm(x, out$mu[1], out$sigma[1]) + 
       out$lambda[2]*dnorm(x, out$mu[2],out$sigma[2])
```

And make a final plot. [Need to move this to ggplot2]


```{r, fig.width=7, fig.height=5}
o.mar <- par("mar")
par(mar=c(4.1,4.1,0.5,0.5))
plot(c(x1, x2), c(0., ymax), type='n', xlab='x', ylab='density')#, 
#    main="Density Estimate of the Gaussian Mixture Model")
points(his$mids, his$density, pch=19, cex=0.75)
lines(truth[, 1], truth[, 2],col="red",lwd=6)
lines(x, fit, col="blue", lwd=2)
legend("topright", c("True Density","Estimated Density", "fit"),
       col=c("black","red", "blue"), lwd=c(NA, 4, 2), pch=c(19,NA,NA))
par(mar=o.mar)
```
