---
title: "Test Gaussian Mixture"
author: "John Minter"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Test Gaussian Mixture}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, cache = F, echo = F, message = F, warning = F, tidy = F, comment=NA}
library(knitr)
options(width = 72)
```

First, let's load the required packages.

```{r load_packages, message=FALSE}
library(particlesizeR)
library(mixtools)
library(ggplot2)
```

Now, let's generate a two Gaussian mixture. Because we consider this a particle size distribution, we need to require the values to be positive.

```{r}
samples <- gen_two_gaussian_mixture(10000, 0.5, 3.0, 1.0, 12.0, 2.0)
samples <- samples[samples > 0]
```

Let's now construct a histogram object. This will give us access to arrays of midpoints and density values.

```{r first_plot, fig.width=7}
library(ggplot2)

df <- data.frame(samples=samples)
plt <- ggplot(df, aes(x=samples)) + geom_histogram(bins=250)
print(plt)
```

Let's look at what is in a ggplot2 histogram, starting at the beginning:

```{r examine_histo_head}
pg <- ggplot_build(plt)
head(pg$data[[1]])
```
Note that we have access to a vector of midpoints (`x`) as well as a vector
of lower boundaries (`xmin`) and a vector of upper boundaries (`xmax`) as
well as the `density` and the `ncount`. 


Next, look at the end of the data...

```{r examine_histo_tail}
tail(pg$data[[1]])
```

Eventually, we want to plot the results. We will build a plot using the newer
`ggplot2` package. We will also generate a series of x (diameter) values to
model.

First, we will generate the data

```{r generate_data}
truth <- gen_two_gaussian_dist(0.5, 3.0, 1.0, 12.0, 2.0, 0, 20, 0.1)

verbose = FALSE

if (verbose == TRUE){
  print(c(min(truth), max(truth)))
  print(max(truth$dist))
  print(max(pg$data[[1]]$density))
  print(head(truth))
}
```

```{r, fig.width=7}
pt_line_size <- 1
df_work <- data.frame(diam=pg$data[[1]]$x, density=pg$data[[1]]$density)
plt <- ggplot(df_work, aes(diam,density)) + 
       geom_point(size=pt_line_size, colour='darkblue' ) +
       geom_line(data=truth, aes(x,dist), colour='darkred', size=pt_line_size) +
       ylab("density") +
       xlab("diameter [nm]") +
       ggtitle("Diameter distribution of two Gaussian distributions") +
       theme(axis.text=element_text(size=12), axis.title=element_text(size=14),
              plot.title = element_text(hjust = 0.5)) + # center the title
      NULL

print(plt)




```


We want to compute the mixture model. We will make a guess at the parameters from the histogram.

```{r mixModel}
out <- normalmixEM(samples, lambda=0.5, mu=c( 2, 9.), sigma=c(.8, 2.))
```

We can now compute the fitted values:

```{r}
x <- seq(from=0, to=20, by=0.1)
fit <- out$lambda[1]*dnorm(x, out$mu[1], out$sigma[1]) + 
       out$lambda[2]*dnorm(x, out$mu[2],out$sigma[2])

df_fit <- data.frame(ecd=x, fit=fit)
```

And make a final plot.

``````{r finalPlot, fig.width=7, fig.height=5}
pt_line_size <- 1
final_plt <- ggplot() +
             geom_point(data=df_work, aes(diam,density),
                        colour="darkblue", size=pt_line_size) +
             geom_line(data=df_fit, aes(ecd, fit),
                       colour="darkred", size=pt_line_size) +
             xlab("size [nm]") +
             ylab("density") +
             scale_x_continuous(breaks = seq(from = 0, to = 20, by = 1),
                               limits = c(0, 20)) +
             scale_y_continuous(breaks = seq(from = 0, to = 0.25, by = 0.05),
                               limits = c(0, 0.25)) +
             ggtitle("Diameter distribution from two Gaussian distributions") +
             labs(caption = 'jrminter@gmail.com / @jrminter') +
             theme(axis.text=element_text(size=12),
                   axis.title=element_text(size=12),
                   plot.title=element_text(hjust = 0.5)) + # center the title
             NULL
print(final_plt)

```
